Locality and Computational Reliability of Linear Response Calculations for Molecular Systems

### 2018-07-12 AnalyticStructure_main.tex
Update on Overleaf.

The response density satisfies an equation of motion written in the form of a quantum Liouville operator (for example see \cite{baroni2008})
\be\lb{LiouvillianRhopomegaDef1}
\left(\omega - \Liouv\right) \dm'_\Phi(\omega) =  \commutator{\op\Phi(\omega)}{\dmnot} \;.
\ee
The Liouvillian superoperator $\Liouv$ for self-consistent systems is expressed as the sum of the unperturbed part plus a coupling term. The action of these 
terms reads, respectively
\be\lb{LiouZeroDef1}
\Liouv \op O \equiv \left(\Liouvnot + \coupl \right) \op O = \commutator{\hnot}{\op O} +\commutator{\op V'[\op O]}{\dmnot}  \;,
\ee
where $\op V'[\op O] \equiv \int \dd \r \dd \r'\frac{\delta \op V[\dmnot]}{\delta \rho(\r,\r')} O(\r,\r')$ encodes the response of the $\dm$-dependent potential to a modification of the density operator.
The inspection of the linear order time evolution of the eigenstates of $\hnot$ evidence that the response density acts as a transition 
operator, linking the occupied and empty subspaces of $\hnot$.
An operator $\op O_\perp$ with this feature satisfy the \emph{transverse} condition  
\be\lb{RhopTransverseDef1}
\op O_\perp \equiv 
\dmnot\op O \op Q_0 + \op Q_0 \op O\dmnot \;,
\ee
where $\op Q_0=\identity-\dmnot$ is the projector in the subspace orthogonal to the occupied states of $\hnot$. 
Thus the condition $\dm'_\Phi = \dm_{\perp}$ holds.

@@

We may then decompose the linear response functional of Eq.~\eqref{LinearResponseFunctDef1}
for any operator $\op O$ as $\braket{\delta\op O}_\Phi=\braket{\delta\op O}^D_\Phi +
\braket{\delta\op O}^C_\Phi$, where
\be
\braket{\delta\op O}^{D,C}_\Phi(\omega) = 
2 \mathrm{Re}\left(\bra{\psi_p} \op O \fscd{D,C} \right)\;.
\ee
For instance, the static polarizability tensor $\alpha_{ij}$ can be easily expressed
from the above quantities as
\be\lb{sminus2}
\alpha_{ij} = \sint \dd a \frac{w^a_{ij}}{\Omega_a^2}\,, \quad w^a_{ij} \equiv \sum_{p\in a} \varphi_p^a[\op r_i]\varphi_p^a[\op r_j]\;.
\ee
Such value can be straightforwardly compared with results of Eq.~\eqref{staticalpha}, where the localized FS is obtained from GS techniques as described in Sec.~\ref{FluctuationState}.
This is the well-know $S_{-2}$ sum-rule\cite{TDDFTBOOK}, that express the value of the static polarizability
in terms of the molecule's oscillator strengths.

@@

\section{Computational details}
The illustrative calculations presented in this paper are all KS-DFT calculations at the LDA level of theory. We employed the BigDFT code \cite{bigdft},
which employs Daubechies Wavelets as computational basis set.
Such orthonormal basis presents optimal features of in view of reproducibility of the results,
and it has proven to represent precisely and explicitly system with Free as well as Periodic boundary conditions (BC), without any basis set superposition error nor supercell aliasing, for Free BC.

>>>>

The response density satisfies an equation of motion written in the form of a quantum Liouville operator (for example see \cite{baroni2008})
\be\lb{LiouvillianRhopomegaDef1}
\left(\omega - \Liouv\right) \dm'_\Phi(\omega) =  \commutator{\op\Phi(\omega)}{\dmnot} \;.
\ee
The Liouvillian superoperator $\Liouv$ for self-consistent systems is expressed as the sum of the unperturbed part $\Liouvnot$ plus a coupling term $ \coupl$. The action of these 
terms reads, respectively
\be\lb{LiouZeroDef1}
\Liouv \op O \equiv \left(\Liouvnot + \coupl \right) \op O = \commutator{\hnot}{\op O} +\commutator{\op V'[\op O]}{\dmnot}  \;,
\ee
where $\op V'[\op O] \equiv \int \dd \r \dd \r'\frac{\delta \op V[\dmnot]}{\delta \rho(\r,\r')} O(\r,\r')$ encodes the response of the $\dm$-dependent potential to a modification of the density operator.
The inspection of the linear order time evolution of the eigenstates of $\hnot$ evidences that the response density acts as a transition 
operator, linking the occupied and empty subspaces of $\hnot$.
An operator $\op O_\perp$ with this feature satisfy the \emph{transverse} condition  
\be\lb{RhopTransverseDef1}
\op O_\perp \equiv 
\dmnot\op O \op Q_0 + \op Q_0 \op O\dmnot \;,
\ee
where $\op Q_0=\identity-\dmnot$ is the projector in the subspace orthogonal to the occupied states of $\hnot$. 
In other terms, the condition $\dm'_\Phi = \dm'_{\Phi\,\perp}$ holds.

@@

We may then decompose the linear response functional of Eq.~\eqref{LinearResponseFunctDef1}
for any operator $\op O$ as $\braket{\delta\op O}_\Phi=\braket{\delta\op O}^D_\Phi +
\braket{\delta\op O}^C_\Phi$, where
\be
\braket{\delta\op O}^{D,C}_\Phi(\omega) = 
2 \mathrm{Re}\left(\bra{\psi_p} \op O \fscd{D,C} \right)\;.
\ee
For instance, the static polarizability tensor $\alpha_{ij}$ can be easily expressed
from the above quantities as
\be\lb{sminus2}
\alpha_{ij} = \sint \dd a \frac{w^a_{ij}}{\Omega_a^2}\,, \quad w^a_{ij} \equiv \sum_{p\in a} \varphi_p^a[\op r_i]\varphi_p^a[\op r_j]\;.
\ee
Such value can be straightforwardly compared with results of Eq.~\eqref{staticalpha}, where the localized FS is obtained from GS techniques as described in Sec.~\ref{FluctuationState}.
This is the well-know $S_{-2}$ sum-rule (see e.g.\cite{Wagner2012}), that express the value of the static polarizability
in terms of the molecule's oscillator strengths.

@@

\section{Computational details}
The illustrative calculations presented in this paper are all KS-DFT calculations at the LDA level of theory. We employed the BigDFT code \cite{BigDFT},
which employs Daubechies Wavelets as computational basis set.
Such orthonormal basis presents optimal features of in view of reproducibility of the results,
and it has proven to represent precisely and explicitly system with Free as well as Periodic boundary conditions (BC), without any basis set superposition error nor supercell aliasing, for Free BC.
